Spectroscopic confirmation of very-wide binaries and large-separation comoving stars from \textsl{Gaia DR1}

### 2017-07-11 paper/comoving-followup.tex

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the pipeline above.
For the remaining 311 pairs, we use the measured radial velocities to re-compute
the marginal likelihood ratio for the two hypotheses considered in previous work
(see Appendix~B, \citealt{Oh:2017}).
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
As with previous work, we set the small dispersion to the orbital velocity of a
binary with the separation of the pair, and use a mixture of three Gaussian
velocity components for the prior, meant to represent the thin disk, thick disk,
and halo velocity distributions relative to the sun.

@@

\todo{Discuss contamination rate vs. separation for ratio cut of 6?}
At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
genuinely comoving; interestingly, the separation distribution of the genuine
pairs appears to be at least uniform or possibly growing in number as the
separation increases.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{dx-dv.pdf}
  \end{center}
  \caption{%
    FML
    \label{fig:dx-dv}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    FML
    \label{fig:separation}}
\end{figure}

@@

\todo{separation distribution - compare with Shaya and Olling sample!}
\todo{10 overlap with Shaya and Olling}
\todo{RAVE pairs - both observed}

@@

\todo{Highlight the MS-Giant pairs}
\todo{Highlight the binary sequence-main sequence pairs}

@@

Abundances!

- Stellar rotation of evolved-main sequence
- Full catalog should contain Shaya and Olling, cross-match to RV surveys

>>>>

Of the 323 observed star pairs, 12 have at least one member with an unreliable
(flagged) radial velocity measurement from the pipeline above.
For the remaining 311 pairs, we use the measured radial velocities to re-compute
the marginal likelihood ratio for the two hypotheses considered in previous work
(see Appendix~B, \citealt{Oh:2017}).
We compute two fully marginalized likelihoods (FMLs) for the data: (1) assuming
that the two stars are comoving with identical 3D velocity vectors (within some
allowed small dispersion), and (2) assuming that the two stars have independent
3D velocity vectors drawn from a prior velocity distribution.
We set the small allowed dispersion ($s$, in previous work) to $1~\kms$, and use
a mixture of three Gaussian velocity components for the prior, meant to
represent the thin disk, thick disk, and halo velocity distributions relative to
the sun.

@@

At separations $\gtrsim 1~\pc$, $\approx 40\%$ of the comoving pairs are
genuinely comoving; interestingly, the separation distribution of the genuine
pairs appears to be uniform or possibly growing in number with increasing
separation.
This strongly suggests the existence of a population of unbound wide binaries or
dissolved open clusters, which maintain approximate spatial coherence for long
times because of the small relative velocities at the time of disruption of the
constituent stars (e.g., \citealt{Jiang:2010}).
The kinematic properties of these unbound binaries and associations contain
information about the tidal field and fluctuations in the mass distribution of
the Milky Way.
With \gaia\ \DR{2}, we will be able to study the statistical properties of these
widely-separated but comoving stars at even larger separations ($>10~\pc$).

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{dx-dv.pdf}
  \end{center}
  \caption{%
    Point estimate of the 3D velocity difference between two stars in a comoving
    pair, $|\bs{v}_1 - \bs{v}_2|$, plotted against a point estimate of the 3D
    spatial separation of the two stars, $|\bs{x}_1 - \bs{x}_2|$.
    The distances used in the point estimates are computed using a
    Lutz-Kelker-corrected parallax value (\citealt{Lutz:1973}).
    \label{fig:dx-dv}}
\end{figure}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist.pdf}
  \end{center}
  \caption{%
    \emph{Left}: Distribution of point estimates of the 3D spatial separation of
    the stars in all observed comoving star pairs (gray), and only those
    identified as genuinely comoving (black). Note that the contamination rate
    is $\approx 60\%$ at large separations ($>1~\pc$), but there are still a
    significant number of genuine comoving pairs at these very wide separations.
    \emph{Right}: Distribution of chord lengths (tangential separations) for all
    comoving star pairs (gray) and only those identified as genuinely comoving
    (black).
    Noise in the parallax measurements makes the 3D separation distribution
    appear roughly uniform.
    Even the distribution in chord length appears to be approximately linear for
    large separations ($\gtrsim 0.1~\pc$).
    \label{fig:separation}}
\end{figure}

@@

@@

We naively expect that comoving, coeval stars should lie along lines parallel to
the main sequence in color-magnitude space. The majority of our star pairs
appear to follow this expectation, with the exception of a few visible outliers.
For example, there appear to be at least a few main-sequence--giant-branch
pairs, and a few pairs that appear to cross the main-sequence.
We highlight a few of these pairs below.

\subsubsection{HD 95088 \& HD 92278}

\todo{Highlight the G subgiant-A pair}

\subsubsection{HD 151090 \& HIP 81988}

\todo{G-subgiant - K-dwarf - well known high proper-motion comoving pair}

\subsubsection{HD 54318 \& TYC 3792-658-1}

\todo{Highlight the largest mag difference MS pair - K-dwarf - B-star}

\subsection{Combined with RAVE}\label{sec:rave}

\begin{figure}[htbp]
  \begin{center}
    \includegraphics[width=\linewidth]{separation-hist-rave.pdf}
  \end{center}
  \caption{%
    Same as \figurename~\ref{fig:separation}, but for our sample of genuine
    comoving stars combined with all genuine comoving stars identified from the
    RAVE overlap.
    \label{fig:separation-RAVE}}
\end{figure}

% \begin{figure}[htbp]
%   \begin{center}
%     \includegraphics[width=\linewidth]{genuine_cmd_rave.pdf}
%   \end{center}
%   \caption{%
%     Color-magnitude diagram of \emph{genuine} comoving star pairs plotted over
%     the density of all \tgas\ sources within 200 pc; similar to
%     \figurename~\ref{fig:sample-cmd}, but for only the genuine comoving
%     pairs.
%     Several interesting pairs that include main-sequence and giant component
%     stars are highlighted in the text
%     (\sectionname~\ref{sec:interesting-pairs}).
%     \label{fig:genuine-cmd}}
% \end{figure}

@@

Abundances - future work, Coronado in prep.! But: throw shade on abundances
because planet engulfment (https://arxiv.org/pdf/1707.02180.pdf). Also Oh et al.
2017b

Dynamical capture at these radii are improbable, but <insert words about binary-binary encounters from Nathan>
